---
title: "EDA"
format: html
editor: visual
---

## Introduction

Briefly describe the data and the variables you have to work with

This data is taken from the Behavioral Risk Factor Surveillance System (BRFSS) which is a annual CDC survey. The data used in this project is from 2015 and contains diabetes health indicators:

-   Diabetes_binary: 0 = no diabetes 1 = prediabetes 2 = diabetes

-   HighBP: 0 = no high BP 1 = high BP

-   HighChol: 0 = no high cholesterol 1 = high cholesterol

-   CholCheck: 0 = no cholesterol check in 5 years 1 = yes cholesterol check in 5 years

-   BMI: Body Mass Index numeric range 12 to 98

-   Smoker: Have you smoked at least 100 cigarettes in your entire life? 0 = no 1 = yes

-   Stroke: (Ever told) you had a stroke 0 = no 1 = yes

-   HeartDiseaseorAttack: coronary heart disease (CHD) or myocardial infarction (MI) 0 = no 1 = yes

-   PhysActivity: physical activity in past 30 days- not including job 0 = no 1 = yes

-   Fruits: Consume Fruit 1 or more times per day 0 = no 1 = yes

-   Veggies: Consume Vegetables 1 or more times per day 0 = no 1 = yes

-   HvyAlcoholConsump: (adult men \>= 14 drinks per week and adult women \>= 7 drinks per week) 0 = no 1 = yes

-   AnyHealthcare: Have any kind of health care coverage, including health insurance, prepaid plans such as HMO, etc. 0 = no 1 = yes

-   NoDocbcCost: Was there a time in the past 12 months when you needed to see a doctor but could not because of cost? 0 = no 1 = yes

-   GenHlth: Would you say that in general your health is: scale 1-5 1 = excellent 2 = very good 3 = good 4 = fair 5 = poor

-   MentHlth: days of poor mental health scale 1-30 days

-   PhysHlth: physical illness or injury days in past 30 days scale 1-30

-   DiffWalk: Do you have serious difficulty walking or climbing stairs? 0 = no 1 = yes

-   Sex: 0 = female 1 = male

-   Age = 13-level age category (\_AGEG5YR see codebook) 1 = 18-24 9 = 60-64 13 = 80 or older

-   Education: Education level (EDUCA see codebook) scale 1-6 1= Never attended school or only kindergarten 2 = elementary etc.

-   Income: Income scale (INCOME2 see codebook) scale 1-8 1 = less than \$10,000 5 = less than \$35,000 8= \$75,000 or more

Describe the purpose of your EDA and ultimate goal of modeling

## Data

```{r}
# load libraries

# relative path to import the data
data <- read.csv("./diabetes_binary_health_indicators_BRFSS2015.csv")

# convert variables to factors with meaningful level names
data$Diabetes_binary <- factor(data$Diabetes_binary,
                           levels= c(0,1,2),
                           labels = c("no diabetes","prediabetes","diabetes"))
data$HighBP <- factor(data$HighBP,
                  levels = c(0,1),
                  labels = c("no high BP", "high BP"))
data$HighChol <- factor(data$HighChol,
                    levels = c(0,1),
                    labels = c("no high cholesterol", "high cholesterol"))
data$CholCheck <- factor(data$CholCheck,
                    levels = c(0,1),
                    labels = c("no cholesterol check in 5 years", "yes cholesterol check in 5 years"))
data <- data |>
  mutate(
    BMI = case_when(
      BMI < 18.5 ~ "underweight",
      BMI >= 18.5 & data$BMI < 24 ~ "healthy weight",
      BMI >= 24 & data$BMI < 30 ~ "overweight",
      BMI >= 30 ~ "obese"
  )
)
data$Smoker <- factor(data$Smoker,
                      levels = c(0,1),
                      labels = c("not a smoker", "smoker"))
data$Stroke <- factor(data$Stroke,
                      levels = c(0,1),
                      labels = c("no", "yes"))
data$HeartDiseaseorAttack <- factor(data$HeartDiseaseorAttack,
                        levels = c(0,1),
                        labels = c("no","yes"))
data$PhysActivity <- factor(data$PhysActivity,
                            levels = c(0,1),
                            labels = c("inactive","active"))
data$Fruits <- factor(data$Fruits,
                      levels = c(0,1),
                      labels = c("no","yes"))
data$Veggies <- factor(data$Veggies,
                       levels = c(0,1),
                       labels = c("no","yes"))
data$HvyAlcoholConsump <- factor(data$HvyAlcoholConsump,
                                 levels = c(0,1),
                                 labels = c("no","yes"))
data$AnyHealthcare <- factor(data$AnyHealthcare,
                             levels = c(0,1),
                             labels = c("no","yes"))
data$NoDocbcCost <- factor(data$NoDocbcCost,
                           levels = c(0,1),
                           labels = c("no","yes"))
data$GenHlth <- factor(data$GenHlth,
                       levels = c(1,2,3,4,5),
                       labels = c("excellent","very good","good","fair","poor"))
data <- data |>
  mutate(
    MentHlth = case_when(
      MentHlth <= 5 ~ "<= 5",
      MentHlth > 5 & MentHlth <= 10 ~ "6-10",
      MentHlth > 10 & MentHlth <= 15 ~ "11-15",
      MentHlth > 15 & MentHlth <= 20 ~ "16-20",
      MentHlth > 20 & MentHlth <= 25 ~ "21-25",
      MentHlth > 25 ~ "26-30"
  )
)
data <- data |>
  mutate(
    PhysHlth = case_when(
      PhysHlth <= 5 ~ "<= 5",
      PhysHlth > 5 & PhysHlth <= 10 ~ "6-10",
      PhysHlth > 10 & PhysHlth <= 15 ~ "11-15",
      PhysHlth > 15 & PhysHlth <= 20 ~ "16-20",
      PhysHlth > 20 & PhysHlth <= 25 ~ "21-25",
      PhysHlth > 25 ~ "26-30"
  )
)
data$DiffWalk <- factor(data$DiffWalk,
                        levels = c(0,1),
                        labels = c("no","yes"))
data$Sex <- factor(data$Sex,
                   levels = c(0,1),
                   labels = c("female","male"))
data$Age <- factor(data$Age,
                   levels = c(1,2,3,4,5,6,7,8,9,10,11,12,13),
                   labels = c("18-24","25-29","30-34","35-39","40-44","45-49","50-54","55-59","60-64","65-69","70-74","75-79","80+"))
data$Education <- factor(data$Education,
                         levels = c(1,2,3,4,5,6),
                         labels = c("Never attended school or only kindergarten","elementary","some high scohol","high school graduate","some college or technical school","college graduate"))
data$Income <- factor(data$Income,
                      levels = c(1,2,3,4,5,6,7,8),
                      labels = c("<$10,000","<$15,000","<$20,000","<$25,000","<$35,000","<$50,000","<$75,000","$75,000+"))

# check on missingness
sum(is.na(data))
# 0 indicates no NAs
```

## Summarizations

```{r}
# meaningful summary statistics and plots about the data (especially as it relates to your response)


# do EDA on full data (not training/testing set first)
```
